---
title: "Exercise1 - Summative Assignment 1 MY472"
author: Raquel Lorenzo Mesa
output: html_document
date: "2 November 2023"
---

## Solution

As indicated in the exercise, we construct an R script defining a function with a single argument 'data', that takes a dataset and performs some input transformation on it. 

For this, we will use a built-in dataset. First, we select it: 
```{r setup}
data()
?UCBAdmissions
```
We choose the dataset with which we will work: 'UCBAdmissions'. This dataset contains aggregate data on applicants to graduate school at Berkeley for the six largest  departments in 1973 classified by admission and sex. It contains 4526 observations on 3 different variables. 
```{r, echo=TRUE}
adminberk <- as.data.frame(UCBAdmissions)
View(adminberk)
```
We defined the function as requested in the exercise.
First, we load the needed library.
```{r}
library(dplyr)
```
In this function, we try to transform the data in a way that serves later to make a further analysis of the information. 

This function takes a dataset as input and has an optional argument named option. The primary purpose of the function is to transform the input dataset into a new dataset with calculated admission percentages. The specific transformation depends on the option provided.

1. When option is set to "admis_percentages," the function groups the data by "Gender" and "Dept" and calculates admission statistics, including the total number of applicants, the total number of admitted applicants, and the admission percentage for each group.

2. *When option is set to "totals_and_percentages," the function groups the data only by "Gender" and calculates similar admission statistics for each gender category.

```{r}
admission_percentages <- function(data, option = "admis_percentages") {
  transformed_data <- data
  if (option == "admis_percentages") {
    transformed_data <- data %>%
      group_by(Gender, Dept) %>%
      summarize(
        Applicants = sum(Freq),
        Admitted = sum(Freq[Admit == "Admitted"]),
        Admission_Percentage = (Admitted / Applicants) * 100
      )
  } else if (option == "totals_and_percentages") {
    transformed_data <- data %>%
      group_by(Gender) %>%
      summarize(
        Applicants = sum(Freq),
        Admitted = sum(Freq[Admit == "Admitted"]),
        Admission_Percentage = (Admitted / Applicants) * 100
      )
  }
  return(transformed_data)
}

```
Then, we can compare the results obtained. 
```{r}
#We use the function in our database
result <- admission_percentages(adminberk)
print(result)

#Then, we will print the solutions with and without the second argument 
#First, we use the option = "admis_percentages". This result is equal to the default one.
result1 <- admission_percentages(adminberk, option = "admis_percentages")
print(result1)

```
As observed, the results from both ways are the same. This means we have established as default option "admis_percentages", so R will choose this option automatically if we do not indicate anything. 

```{r}
#Second, we use the option = "totals_and_percentages".
#This will show the total number of applicants and admitted people in the university,
#only by gender 
result2 <- admission_percentages(adminberk, option = "totals_and_percentages")
print(result2)
```
As observed, the results change when we set the second argument option equal to "totals_and_percentages", as already explained previously. 

As an interesting fact, this database serves as a good example of the Simpson's Paradox. When looking at the data as a whole (result2), we can observe a positive bias against men. However, when further analyzing the data, we can see how this bias disappears if we calculate the different percentages for each specific department of the university. 
